%% Copyright 2006-2013 Xavier Danaux (xdanaux@gmail.com).
%
% This work may be distributed and/or modified under the
% conditions of the LaTeX Project Public License version 1.3c,
% available at http://www.latex-project.org/lppl/.


\documentclass[11pt,a4paper,sans]{moderncv}        % possible options include font size ('10pt', '11pt' and '12pt'), paper size ('a4paper', 'letterpaper', 'a5paper', 'legalpaper', 'executivepaper' and 'landscape') and font family ('sans' and 'roman')

\usepackage{tabularx}

% modern themes
\moderncvstyle{banking}                            % style options are 'casual' (default), 'classic', 'oldstyle' and 'banking'
\moderncvcolor{blue}                                % color options 'blue' (default), 'orange', 'green', 'red', 'purple', 'grey' and 'black'
%\renewcommand{\familydefault}{\sfdefault}         % to set the default font; use '\sfdefault' for the default sans serif font, '\rmdefault' for the default roman one, or any tex font name
%\nopagenumbers{}                                  % uncomment to suppress automatic page numbering for CVs longer than one page

% character encoding
\usepackage[utf8]{inputenc}                       % if you are not using xelatex ou lualatex, replace by the encoding you are using
%\usepackage{CJKutf8}                              % if you need to use CJK to typeset your resume in Chinese, Japanese or Korean

% adjust the page margins
\usepackage[scale=0.75]{geometry}
%\setlength{\hintscolumnwidth}{3cm}                % if you want to change the width of the column with the dates
%\setlength{\makecvtitlenamewidth}{10cm}           % for the 'classic' style, if you want to force the width allocated to your name and avoid line breaks. be careful though, the length is normally calculated to avoid any overlap with your personal info; use this at your own typographical risks...

\usepackage{import}

% personal data
\name{Evan `Pete'}{Walsh}
\title{CV}                               % optional, remove / comment the line if not wanted
\address{Iowa State University}{Department of Statistics}{Snedecor Hall Room 2410, Ames, IA 50011} % optional, remove / comment the line if not wanted; the "postcode city" and and "country" arguments can be omitted or provided empty
% \phone[mobile]{+44 12345 12345}                   % optional, remove / comment the line if not wanted
% \phone[fixed]{01234 123456}                    % optional, remove / comment the line if not wanted
%\phone[fax]{+3~(456)~789~012}                      % optional, remove / comment the line if not wanted
\email{epwalsh@iastate.edu}                               % optional, remove / comment the line if not wanted
\homepage{epwalsh.com}                         % optional, remove / comment the line if not wanted
\extrainfo{\href{https://github.com/epwalsh}{\faGithub}, \href{https://www.linkedin.com/in/epwalsh}{\faLinkedin}: epwalsh} % optional, remove / comment the line if not wanted
%\photo[64pt][0.4pt]{picture}                       % optional, remove / comment the line if not wanted; '64pt' is the height the picture must be resized to, 0.4pt is the thickness of the frame around it (put it to 0pt for no frame) and 'picture' is the name of the picture file
%\quote{Some quote}                                 % optional, remove / comment the line if not wanted

% to show numerical labels in the bibliography (default is to show no labels); only useful if you make citations in your resume
%\makeatletter
%\renewcommand*{\bibliographyitemlabel}{\@biblabel{\arabic{enumiv}}}
%\makeatother
%\renewcommand*{\bibliographyitemlabel}{[\arabic{enumiv}]}% CONSIDER REPLACING THE ABOVE BY THIS

% bibliography with mutiple entries
%\usepackage{multibib}
%\newcites{book,misc}{{Books},{Others}}


%----------------------------------------------------------------------------------
%            content
%----------------------------------------------------------------------------------
\begin{document}
%\begin{CJK*}{UTF8}{gbsn}                          % to typeset your resume in Chinese using CJK
%-----       resume       ---------------------------------------------------------
\makecvtitle

\small{Third-year graduate student pursuing a co-major PhD in Statistics and Mathematics.}

\section{Education}

\vspace{5pt}

\begin{itemize}

  \item{\cventry{August 2014 -- Present}{Pursuing PhD in Statistics and Mathematics (co-major)}{Iowa State University}{Ames, IA}{}{Overall GPA: 3.86 / 4.00}}

  \item{\cventry{August 2010 -- May 2014}{Bachelor of Science (B.S.), Mathematics (with honors)}{St. Lawrence University}{Canton, NY}{}{Major GPA:
    3.95 / 4.00}}  

  \item{\cventry{Spring 2013}{Semester Abroad, Mandarin Chinese and Economics}{East China Normal University}{Shanghai, China}{}{}}

\end{itemize}

\section{Academic Experience}

\vspace{6pt}

\subsection{Iowa State University}

\begin{itemize}

  \item{\cventry{Spring 2016 -- Present}{AFLEX}{Research Assistant}{}{}{\vspace{3pt} An interdisciplinary approach using machine learning and 
        natural language processing to develop an Automated Functional Language EXtraction (AFLEX) system to transform the translation of STEM 
    research to society; funded by NIH, working with Dr. Kris De Brabanter.}}

    \vspace{6pt}

  \item{\cventry{Spring 2016}{STAT 105}{Instructor}{}{}{\vspace{3pt}Intro Stat for Engineering Students (online section)}}

    \vspace{6pt}

  \item{\cventry{Spring 2015}{STAT 101}{Instructor}{}{}{\vspace{3pt}Principles of Statistics}}

    \vspace{6pt}

  \item{\cventry{Fall 2014, Fall 2015}{STAT 326}{Lab Instructor}{}{}{\vspace{3pt}Business Statistics II}}

    \vspace{6pt}

  \item{\cventry{Fall 2014}{STAT 101}{Lab Instructor}{}{}{\vspace{3pt}Principles of Statistics}}

\end{itemize}

\subsection{St. Lawrence University}

\begin{itemize}

  \item{\cventry{Fall 2013 -- Spring 2014}{Analysis of River Data}{Research Assistant}{}{}{\vspace{3pt} Assisted the Chemistry Department in analyzing and visualizing data using R.}}

    \vspace{6pt}

  \item{\cventry{Fall 2012 -- Spring 2014}{Quantitative Resource Center}{Peer Mentor}{}{}{\vspace{3pt}Tutored other students and occasionally ran
    workshops in mathematics, computer science, statistics, and economics.}}

\end{itemize}

\section{Honors and Awards}

\vspace{6pt}

\subsection{Iowa State University}

\begin{itemize}
  \item \textbf{2nd Place - prudsys 2015 Data Mining Cup} \hfill \emph{Spring 2015} \\ 
    The Data Mining Cup, hosted by prudsys AG, is an international data
    mining competition for universities. In 2015, a total of 188 teams from 48
    countries participated in the challenge which involved predicting coupon
    usage rates and basket values for real customers of an online shop. Our
    predictions landed us 2nd overall and we were invited to Berlin to
    present our results.
\end{itemize}

\vspace{6pt}

\subsection{St. Lawrence University}

\begin{itemize}
  \item \textbf{Honors in Mathematics} \hfill \emph{May 2014}

    \vspace{6pt}

  \item \textbf{Magna Cum Laude} \hfill \emph{May 2014}

    \vspace{6pt}

  \item \textbf{NCAA Liberty League All-Academic Team} \hfill \emph{Fall 2010, Fall 2011, Fall 2012}

\end{itemize}

\section{Computer Skills}

\vspace{6pt}

\begin{itemize}
  \item \textbf{General Programming:}  Competent with Python; some experience with C, C++, and Java. 

\vspace{6pt}

\item \textbf{Data Specific Programming:} Proficient with R and Python (pandas, sklearn, numpy, etc.).

\vspace{6pt}

\item \textbf{Presentation / Visualization:} Experienced with \LaTeX, HTML5, CSS3, and various R libraries such Shiny, ggplot2, ggvis, and RMarkdown.

\vspace{6pt}

\item \textbf{Workflow:} Frequently use Git, Vim, and other command line tools; knowledge of basic Bash scripting.

\vspace{6pt}

\item \textbf{Operating Systems:} Most comfortable in a Unix or Linux environment.

\end{itemize}


\section{Coursework}

\vspace{6pt}

\subsection{Iowa State University}

\begin{tabularx}{\textwidth}{lcXl}
  \textbf{Department} & \ \ \textbf{\#} \ \ & \textbf{Description} & \textbf{Semester} \\
  \\
  % \hline
  MATH/STAT & 642 & Advanced probability theory & Spring 2016 \\
  MATH & 516 & Real analysis II & - \\
  MATH & 511 & Theory of functions of a single complex variable  & - \\
  \\
  % \hline
  MATH/STAT & 641 & Foundations of measure and probability theory & Fall 2015 \\
  MATH & 515 & Real analysis I & - \\
  \\
  % \hline
  MATH & 501 & An introduction to real analysis & Summer 2015 \\
  \\
  % \hline
  STAT & 580 & Statistical computing with C (audit) & Spring 2015 \\
  CS & 573 & An introduction to machine learning & - \\
  STAT & 543 & Theory of probability and statistics II & - \\
  STAT & 510 & Statistical methods II & - \\
  \\
  % \hline
  STAT & 579 & Statistical computing with R & Fall 2014 \\
  STAT & 542 & Theory of probability and statistics I & - \\
  STAT & 500 & Statistical methods I & - \\
  % \hline
\end{tabularx}


\end{document}
