---
title: "Unit 1 Topics"
runtime: shiny
output: html_document
---

### Misleading histograms

> The way data is represented can potentially be misleading. Below is a histogram of a nice unimodal, bell-shaped, symmetric distribution... Right? Wrong. Adjust the binwidths and find out.

```{r echo=FALSE}
source("hist_widget.R")
hist_widget()
```

### The effect of skewness on mean and median

> This plot is meant to illustrate how the skewness of a distribution changes the relationship between the mean and the median. The mean is represented by the blue dashed line, while the median is the green dashed line.

```{r echo=FALSE}
source("skew_widget.R")
skew_widget()
```

### How limits on the range of the data can affect shape

> Data that have a natural limit to their range, for example grades (between 0 and 100), the shape of the distribution will depend on how close the center is to these limits and how big the variability is.

```{r echo=FALSE}
source("limit_widget.R")
limit_widget()
```

