---
title: "Chapter 5"
author: "The Standard Deviation as a Ruler and the Normal Model"
output: 
  ioslides_presentation:
    logo: ISU.png
---

## Chapter Outline

- Standardizing observations for comparisons
- The normal model
- 68-95-99.7 rule for the normal model
- Finding percentiles and areas for given observations from a normal model
- Finding observations for given percentiles or areas from a normal model

# Standardizing observations

## Comparing two variables

- ACT score
- SAT score
- Different scale
- Measuring similar values
- How do you compare them?

## Example 1

- The average score on the ACT English exam is 21.0, with a standard deviation of 4.0. The average score on the SAT verbal exam is 520 with a standard deviation of 100.
    - Albert scores a 27 on the ACT English exam
    - Alfred scores a 770 on the SAT Verbal exam
- Who has the better score? Albert or Alfred?

## Standardizing observations

- $y =$ observation of quantitative variable
- How does the value of $y$ relate to the mean value?
- How does the value of $y$ for this quantitative variable relate to another observation of a different quantitative variable?

## Standardizing variables

$$z = \frac{y - \bar{y}}{s}$$

- $z$ has no units (just a number)
- Puts observations on same scale
    - Mean (center) at 0
    - Standard deviation (spread) of 1
- Does not change overall shape of the distribution

## Standardizing variables

- $z =$ # of standard deviations observation is away from mean
    - Negative $z$: observation is below mean
        - Ex: $z=-1.5\Rightarrow$ observation is 1.5 standard devations below mean
    - Positive $z$: observation is above mean
        - Ex: $z=0.5\Rightarrow$ observation is 0.5 standard deviations above mean

## Recall: Example 1

- The average score on the ACT English exam is 21.0, with a standard deviation of 4.0. The average score on the SAT verbal exam is 520 with a standard deviation of 100.
    - Albert scores a 27 on the ACT English exam
    - Alfred scores a 770 on the SAT Verbal exam
- How do we answer who has the better score?
    - Compute z-scores and compare
    
## Example 1

- Albert $$z = \frac{y_{1} - \bar{y}}{s} = \frac{27-21}{4} = 1.5$$
- Alfred $$z = \frac{y_{2} - \bar{y}}{s} = \frac{770 - 520}{100} = 2.5$$
- Who has the better scores?
    
## Example 2

- The average score on the ACT Math exam is 20.7 with a standard deviation of 4.1. The average score on the SAT Math exam is 510 with a standard deviation of 100
    - Alberta scores a 15 on the ACT Math exam
    - Fredda scores a 340 on the SAT Math exam
- Who has the better score? Alberta or Fredda?

## Example 2

- Alberta $$z = \frac{y_{1} - \bar{y}}{s} = \frac{15-20.7}{4.1} = -1.39$$
- Fredda $$z = \frac{y_{2} - \bar{y}}{s} = \frac{340 - 510}{100} = -1.70$$
- Who has the better score?

## Your turn

[In-class worksheet](http://epwalsh.github.io/Stat101/Topics/Unit-2/z-scores.worksheet.pdf)
    
## Distributions and standardizing

- Standardizing
    - Allows you to make comparisons of observations between different variables
    - Without the distribution information, you still don't know anything about the percentile value of your observation
    - This percentile value depends on the distribution
    
## Models for data distributions

- Different models according to different aspects of distributions
    - Shape
    - Center
    - Variability
    
# The Normal Model

## Normal Model

- Shape
    - Unimodal
    - Symmetric
    - Bell-shaped
- Determined by two parameters
    - Mean $(\mu)$
    - Standard deviation $(\sigma)$
    
## Connection to data

- No data distribution follows a normal model exactly
- Many data distributions are very close though
- How do you know?
    - Histogram
    - Normal Quantile Plot (will learn in lab)
    
## Example

```{r echo=FALSE, message=F}
library(ggplot2)
set.seed(1)
fHeights <- rnorm(150, 66.5, 2)
N <- length(fHeights)
bw <- 1
dnorm.count <- function(x, mean = 0, sd = 1, log = FALSE, n = 1, binwidth = 1) {
  n * binwidth * dnorm(x = x, mean = mean, sd = sd, log = log)
}
female_height <- ggplot(data.frame(fHeights), aes(fHeights)) + xlab("Height") + ylab("Count") + 
  geom_histogram(aes(y=..count..), binwidth=bw, colour="black", fill="#333333") 
female_height + ggtitle("Distribution of Female Heights")
```

## Example

```{r echo=FALSE, message=FALSE}
female_height +
  stat_function(geom="line", fun=dnorm.count, colour=("green"), size=1.5,
                arg=list(mean=66.5, sd=2, binwidth=bw, n = N)) + 
  ggtitle("Distribution of Female Heights with Normal Model Overlayed")
```

## A closer look at the normal model

- Mean $\mu$ 
    - Locates the center of the distribution
    - Splits the curve in half
- Standard deviation $\sigma$
    - Controls the variability of curve
    - Ruler of distribution
- Write as $N(\mu, \sigma)$

## Example

- Height of men is normally distributed 
    - Mean $\mu = 70$ inches
    - Standard deviation $\sigma = 3$ inches
- So, the height of men is distributed $N(70,3)$
- What percent of men have heights less than 70 inches? 
- What percent of men have heights greater than 70 inches?
    
## Example - Height of men

```{r echo=FALSE}
x <- seq(60, 80, by=0.05)
y <- dnorm(x,70,3)
mHeights <- data.frame(x,y)
male_height <- qplot(x,y, mHeights, geom="line") + xlab("Male Height") + ylab("") 
male_height
```

# 68-95-99.7 rule

## 68-95-99.7 rule

- Approximately 68% of observations are within 1 $\sigma$ of the mean $\mu$
- If $\mu = 0$ and $\sigma = 1$, approximately 68% of the observations are between -1 and 1

## 68-95-99.7 rule

```{r echo=FALSE}
x1 <- seq(-4,4,by=0.01)
y1 <- dnorm(x1)
normDF <- data.frame(x1,y1)
plotNorm68 <- qplot(x1,y1,normDF, geom="line") + 
  geom_ribbon(data=subset(normDF, x1 >= -1 & x1 <= 1), aes(ymax=y1), ymin=0, fill="red",
              colour=NA, alpha=0.5) +
  geom_segment(aes(x=c(-1,1), y=c(0,0), xend=c(-1,1), yend=c(dnorm(-1),dnorm(1)))) +
  xlab("z") + ylab("") + annotate("text", x=0, y=0.15, label="68%", size=14)
plotNorm68
```

## 68-95-99.7 rule

- Approximately 95% of observations are within $2\sigma$ of the mean $\mu$
- If $\mu = 0$ and $\sigma = 1$, approximately 95% of the observations are between $-2$ and $2$

## 68-95-99.7 rule

```{r echo=FALSE}
plotNorm95 <- qplot(x1,y1,normDF, geom="line") + 
  geom_ribbon(data=subset(normDF, x1 >= -2 & x1 <= 2), aes(ymax=y1), ymin=0, fill="red",
              colour=NA, alpha=0.5) +
  geom_segment(aes(x=c(-2,2), y=c(0,0), xend=c(-2,2), yend=c(dnorm(-2),dnorm(2)))) +
  xlab("z") + ylab("") + annotate("text", x=0, y=0.15, label="95%", size=14)
plotNorm95
```

## 68-95-99.7 rule

- Approximately 99.7% of observations are within $3\sigma$ of the mean $\mu$
- If $\mu = 0$ and $\sigma = 1$, approximately 99.7% of the observations are between $-3$ and $3$

## 68-95-99.7 rule

```{r echo=FALSE}
plotNorm99 <- qplot(x1,y1,normDF, geom="line") + 
  geom_ribbon(data=subset(normDF, x1 >= -3 & x1 <= 3), aes(ymax=y1), ymin=0, fill="red",
              colour=NA, alpha=0.5) +
  geom_segment(aes(x=c(-3,3), y=c(0,0), xend=c(-3,3), yend=c(dnorm(-3),dnorm(3)))) +
  xlab("z") + ylab("") + annotate("text", x=0, y=0.15, label="99.7%", size=14)
plotNorm99
```

## Example - Height of men

- Height of men $\sim N(70, 3)$
- 68-95-99.7 rule
    - 68% of men will have heights between which two values?
        - $\mu-\sigma = 70 - 3 = 67$ inches
        - $\mu+\sigma = 70 + 3 = 73$ inches

```{r echo=FALSE, eval=FALSE}
male_height + geom_segment(aes(x=61, y=0, xend=61, yend=dnorm(61,70,3))) +
  geom_segment(aes(x=64, y=0, xend=64, yend=dnorm(64,70,3))) +
  geom_segment(aes(x=67, y=0, xend=67, yend=dnorm(67,70,3))) +
  geom_segment(aes(x=73, y=0, xend=73, yend=dnorm(73,70,3))) +
  geom_segment(aes(x=76, y=0, xend=76, yend=dnorm(76,70,3))) +
  geom_segment(aes(x=79, y=0, xend=79, yend=dnorm(79,70,3))) +
  geom_ribbon(data=subset(mHeights, x >= 61 & x <= 79), aes(ymax=y), ymin=0, fill="red",
              colour=NA, alpha=0.2) +
  geom_ribbon(data=subset(mHeights, x >= 64 & x <= 76), aes(ymax=y), ymin=0, fill="red",
              colour=NA, alpha=0.3) +
  geom_ribbon(data=subset(mHeights, x >= 67 & x <= 73), aes(ymax=y), ymin=0, fill="red",
              colour=NA, alpha=0.3)
```

## 68% rule implies

```{r echo=FALSE}
male_height + 
  geom_ribbon(data=subset(mHeights, x >= 67 & x <= 73), aes(ymax=y), ymin=0, fill="red",
              colour=NA, alpha=0.5) +
  geom_segment(aes(x=67, y=0, xend=67, yend=dnorm(67,70,3))) +
  geom_segment(aes(x=73, y=0, xend=73, yend=dnorm(73,70,3))) +
  annotate("text", x=70, y=0.05, label="68%", size=14)
```

## 68% rule implies

```{r echo=FALSE}
male_height + 
  geom_ribbon(data=subset(mHeights, x >= 67 & x <= 70), aes(ymax=y), ymin=0, fill="red",
              colour=NA, alpha=0.5) +
  geom_segment(aes(x=67, y=0, xend=67, yend=dnorm(67,70,3))) +
  geom_segment(aes(x=70, y=0, xend=70, yend=dnorm(70,70,3))) +
  annotate("text", x=68.5, y=0.05, label="?%", size=8)
```

## 68% rule implies

```{r echo=FALSE}
male_height + 
  geom_ribbon(data=subset(mHeights, x >= 70 & x <= 73), aes(ymax=y), ymin=0, fill="red",
              colour=NA, alpha=0.5) +
  geom_segment(aes(x=73, y=0, xend=73, yend=dnorm(73,70,3))) +
  geom_segment(aes(x=70, y=0, xend=70, yend=dnorm(70,70,3))) +
  annotate("text", x=71.5, y=0.05, label="?%", size=8)
```

## What percent is below 1 standard deviation below the mean?

```{r echo=FALSE}
male_height + 
  geom_ribbon(data=subset(mHeights, x >= 60 & x <= 67), aes(ymax=y), ymin=0, fill="red",
              colour=NA, alpha=0.5) +
  geom_segment(aes(x=67, y=0, xend=67, yend=dnorm(67,70,3))) 
```

## What percent is below 1 standard deviation below the mean?

```{r echo=FALSE}
male_height + 
  geom_ribbon(data=subset(mHeights, x >= 60 & x <= 67), aes(ymax=y), ymin=0, fill="red",
              colour=NA, alpha=0.5) +
  geom_segment(aes(x=67, y=0, xend=67, yend=dnorm(67,70,3))) +
  annotate("text", x=62.5, y=0.06, label="(100-68)/2\n=16%", size=8)
```

## What percent is above 1 standard deviation below the mean?

```{r echo=FALSE}
male_height + 
  geom_ribbon(data=subset(mHeights, x >= 67 & x <= 80), aes(ymax=y), ymin=0, fill="red",
              colour=NA, alpha=0.5) +
  geom_segment(aes(x=67, y=0, xend=67, yend=dnorm(67,70,3))) 
```

## What percent is above 1 standard deviation below the mean?

```{r echo=FALSE}
male_height + 
  geom_ribbon(data=subset(mHeights, x >= 67 & x <= 80), aes(ymax=y), ymin=0, fill="red",
              colour=NA, alpha=0.5) +
  geom_segment(aes(x=67, y=0, xend=67, yend=dnorm(67,70,3))) +
  annotate("text", x=77.4, y=0.06, label="(100-16)=84%", size=8)
```

## What percent is above 1 standard deviation above the mean?

```{r echo=FALSE}
male_height + 
  geom_ribbon(data=subset(mHeights, x >= 73 & x <= 80), aes(ymax=y), ymin=0, fill="red",
              colour=NA, alpha=0.5) +
  geom_segment(aes(x=73, y=0, xend=73, yend=dnorm(73,70,3))) 
```

## What percent is above 1 standard deviation above the mean?

```{r echo=FALSE}
male_height + 
  geom_ribbon(data=subset(mHeights, x >= 73 & x <= 80), aes(ymax=y), ymin=0, fill="red",
              colour=NA, alpha=0.5) +
  geom_segment(aes(x=73, y=0, xend=73, yend=dnorm(73,70,3))) +
  annotate("text", x=77.5, y=0.06, label="(100-68)/2\n=16%", size=8)
```

## What percent is below 1 standard deviation above the mean?

```{r echo=FALSE}
male_height + 
  geom_ribbon(data=subset(mHeights, x >= 60 & x <= 73), aes(ymax=y), ymin=0, fill="red",
              colour=NA, alpha=0.5) +
  geom_segment(aes(x=73, y=0, xend=73, yend=dnorm(73,70,3))) 
```

## What percent is below 1 standard deviation above the mean?

```{r echo=FALSE}
male_height + 
  geom_ribbon(data=subset(mHeights, x >= 60 & x <= 73), aes(ymax=y), ymin=0, fill="red",
              colour=NA, alpha=0.5) +
  geom_segment(aes(x=73, y=0, xend=73, yend=dnorm(73,70,3))) +
  annotate("text", x=62.6, y=0.06, label="(100-16)=84%", size=8)
```

## Example

- Height of men $\sim N(70,3)$
- 68-95-99.7 rule
    - 95% of men will have heights between which two values?
        
## Example

- Height of men $\sim N(70,3)$
- 68-95-99.7 rule
    - 99.7% of men will have heights between which two values?

## Beyond the 68-95-99.7 rule

- Z-table
    - Connection between normal values and percentiles for standard normal model
- Standard normal model
    - Denoted by $Z$
    - Mean $\mu = 0$
    - Standard deviation $\sigma = 1$
    
# Finding percentiles and areas for given observations from a normal model
    
## Z Table

- Table gives proportion of curve below a particular $z$ score (the percentile for the value $z$)
    - $z$ values range from $-3.90$ to $3.90$ on the table
    - Row - ones and tenths place for $z$
    - Column - hundredths place for $z$
    
## Percentile for z = -1.50

```{r echo=FALSE}
qplot(x1,y1,normDF, geom="line") + 
  geom_ribbon(data=subset(normDF, x1 >= -4 & x1 <= -1.50), aes(ymax=y1), ymin=0, fill="red",
              colour=NA, alpha=0.5) +
  geom_segment(aes(x=-1.5, y=0, xend=-1.5, yend=dnorm(-1.50))) +
  xlab("z") + ylab("") + ggtitle("Standard Normal Distribution")
  # + annotate("text", x=0, y=0.15, label="95%", size=14)
```

## Percentile for z = -1.50

```{r echo=FALSE}
qplot(x1,y1,normDF, geom="line") + 
  geom_ribbon(data=subset(normDF, x1 >= -4 & x1 <= -1.50), aes(ymax=y1), ymin=0, fill="red",
              colour=NA, alpha=0.5) +
  geom_segment(aes(x=-1.5, y=0, xend=-1.5, yend=dnorm(-1.50))) +
  xlab("z") + ylab("") + ggtitle("Standard Normal Distribution") + 
  annotate("text", x=-2.5, y=0.2, label="P(z < -1.50)\n=0.0668", size=8)
```

## Percentile for z = 1.98

```{r echo=FALSE}
qplot(x1,y1,normDF, geom="line") + 
  geom_ribbon(data=subset(normDF, x1 >= -4 & x1 <= 1.98), aes(ymax=y1), ymin=0, fill="red",
              colour=NA, alpha=0.5) +
  geom_segment(aes(x=1.98, y=0, xend=1.98, yend=dnorm(1.98))) +
  xlab("z") + ylab("") + ggtitle("Standard Normal Distribution")
```

## Percentile for z = 1.98

```{r echo=FALSE}
qplot(x1,y1,normDF, geom="line") + 
  geom_ribbon(data=subset(normDF, x1 >= -4 & x1 <= 1.98), aes(ymax=y1), ymin=0, fill="red",
              colour=NA, alpha=0.5) +
  geom_segment(aes(x=1.98, y=0, xend=1.98, yend=dnorm(1.98))) +
  xlab("z") + ylab("") + ggtitle("Standard Normal Distribution") +
  annotate("text", x=-2.5, y=0.2, label="P(z < 1.98)\n=0.9761", size=8)
```

## Proportion greater than z = -1.65

```{r echo=FALSE}
qplot(x1,y1,normDF, geom="line") + 
  geom_ribbon(data=subset(normDF, x1 >= -1.65), aes(ymax=y1), ymin=0, fill="red",
              colour=NA, alpha=0.5) +
  geom_segment(aes(x=-1.65, y=0, xend=-1.65, yend=dnorm(-1.65))) +
  xlab("z") + ylab("") + ggtitle("Standard Normal Distribution")
```

## Proportion greater than z = -1.65

```{r echo=FALSE}
qplot(x1,y1,normDF, geom="line") + 
  geom_ribbon(data=subset(normDF, x1 >= -1.65), aes(ymax=y1), ymin=0, fill="red",
              colour=NA, alpha=0.5) +
  geom_segment(aes(x=-1.65, y=0, xend=-1.65, yend=dnorm(-1.65))) +
  xlab("z") + ylab("") + ggtitle("Standard Normal Distribution") +
  annotate("text", x=2.5, y=0.3, 
           label="P(z > -1.65)\n= 1 - P(z < -1.65)\n= 1 - 0.0495\n= 0.9761", size=8)
```

## Proportion greater than z = 0.73

```{r echo=FALSE}
qplot(x1,y1,normDF, geom="line") + 
  geom_ribbon(data=subset(normDF, x1 >= 0.73), aes(ymax=y1), ymin=0, fill="red",
              colour=NA, alpha=0.5) +
  geom_segment(aes(x=0.73, y=0, xend=0.73, yend=dnorm(0.73))) +
  xlab("z") + ylab("") + ggtitle("Standard Normal Distribution")
```

## Proportion greater than z = 0.73

```{r echo=FALSE}
qplot(x1,y1,normDF, geom="line") + 
  geom_ribbon(data=subset(normDF, x1 >= 0.73), aes(ymax=y1), ymin=0, fill="red",
              colour=NA, alpha=0.5) +
  geom_segment(aes(x=0.73, y=0, xend=0.73, yend=dnorm(0.73))) +
  xlab("z") + ylab("") + ggtitle("Standard Normal Distribution") + 
  annotate("text", x=2.5, y=0.3, 
           label="P(z > 0.73)\n= 1 - P(z < 0.73)\n= 1 - 0.7673\n= 0.0.2327", size=8)
```

## Proportion between z = 0.5 and z = 1.4

```{r echo=FALSE}
qplot(x1,y1,normDF, geom="line") + 
  geom_ribbon(data=subset(normDF, x1 >= 0.5 & x1 <= 1.4), aes(ymax=y1), ymin=0, fill="red",
              colour=NA, alpha=0.5) +
  geom_segment(aes(x=0.5, y=0, xend=0.5, yend=dnorm(0.5))) +
  geom_segment(aes(x=1.4, y=0, xend=1.4, yend=dnorm(1.4))) +
  xlab("z") + ylab("") + ggtitle("Standard Normal Distribution")
```

## Proportion between z = 0.5 and z = 1.4

```{r echo=FALSE}
qplot(x1,y1,normDF, geom="line") + 
  geom_ribbon(data=subset(normDF, x1 >= 0.5 & x1 <= 1.4), aes(ymax=y1), ymin=0, fill="red",
              colour=NA, alpha=0.5) +
  geom_segment(aes(x=0.5, y=0, xend=0.5, yend=dnorm(0.5))) +
  geom_segment(aes(x=1.4, y=0, xend=1.4, yend=dnorm(1.4))) +
  xlab("z") + ylab("") + ggtitle("Standard Normal Distribution") +
  annotate("text", x=2.5, y=0.3, 
           label="P(0.5<z<1.4)\n=P(z<1.4)-P(z<0.5)\n=0.9192-0.6915\n=0.2277", size=8)
```

## Proportion between z = -2.3 and z = -0.05

```{r echo=FALSE}
qplot(x1,y1,normDF, geom="line") + 
  geom_ribbon(data=subset(normDF, x1 >= -2.3 & x1 <= -0.05), aes(ymax=y1), ymin=0, fill="red",
              colour=NA, alpha=0.5) +
  geom_segment(aes(x=-2.3, y=0, xend=-2.3, yend=dnorm(-2.3))) +
  geom_segment(aes(x=-0.05, y=0, xend=-0.05, yend=dnorm(-0.05))) +
  xlab("z") + ylab("") + ggtitle("Standard Normal Distribution")
```

## Proportion between z = -2.3 and z = -0.05

```{r echo=FALSE}
qplot(x1,y1,normDF, geom="line") + 
  geom_ribbon(data=subset(normDF, x1 >= -2.3 & x1 <= -0.05), aes(ymax=y1), ymin=0, fill="red",
              colour=NA, alpha=0.5) +
  geom_segment(aes(x=-2.3, y=0, xend=-2.3, yend=dnorm(-2.3))) +
  geom_segment(aes(x=-0.05, y=0, xend=-0.05, yend=dnorm(-0.05))) +
  xlab("z") + ylab("") + ggtitle("Standard Normal Distribution") +
  annotate("text", x=2.5, y=0.3, 
           label="P(-2.3<z<-0.05)\n=P(z<-0.05)-P(z<-2.3)\n=0.4801-0.0107\n=0.0.4694", size=7)
```


# Finding z from a given percentile

## Finding z from a given percentile

```{r echo=FALSE}
qplot(x1,y1,normDF, geom="line") + 
  geom_ribbon(data=subset(normDF, x1 <= -0.15), aes(ymax=y1), ymin=0, fill="blue",
              colour=NA, alpha=0.5) +
  geom_segment(aes(x=-0.15, y=0, xend=-0.15, yend=dnorm(-0.15))) +
  xlab("z") + ylab("") + ggtitle("Standard Normal Distribution") +
  annotate("text", x=-1, y=0.1, label="0.4404", size=7)
```

## Finding z from a given percentile

```{r echo=FALSE}
qplot(x1,y1,normDF, geom="line") + 
  geom_ribbon(data=subset(normDF, x1 <= -0.15), aes(ymax=y1), ymin=0, fill="blue",
              colour=NA, alpha=0.5) +
  geom_segment(aes(x=-0.15, y=0, xend=-0.15, yend=dnorm(-0.15))) +
  xlab("z") + ylab("") + ggtitle("Standard Normal Distribution") +
  annotate("text", x=-1, y=0.1, label="0.4404", size=7) +
  annotate("text", x=-3, y=0.2, label="z = -0.15", size =8)
```

## Finding z from a given percentile

```{r echo=FALSE}
qplot(x1,y1,normDF, geom="line") + 
  geom_ribbon(data=subset(normDF, x1 <= 2.13), aes(ymax=y1), ymin=0, fill="blue",
              colour=NA, alpha=0.5) +
  geom_segment(aes(x=2.13, y=0, xend=2.13, yend=dnorm(2.13))) +
  xlab("z") + ylab("") + ggtitle("Standard Normal Distribution") +
  annotate("text", x=0, y=0.1, label="0.9834", size=8)
```

## Finding z from a given percentile

```{r echo=FALSE}
qplot(x1,y1,normDF, geom="line") + 
  geom_ribbon(data=subset(normDF, x1 <= 2.13), aes(ymax=y1), ymin=0, fill="blue",
              colour=NA, alpha=0.5) +
  geom_segment(aes(x=2.13, y=0, xend=2.13, yend=dnorm(2.13))) +
  xlab("z") + ylab("") + ggtitle("Standard Normal Distribution") +
  annotate("text", x=0, y=0.1, label="0.9834", size=8) +
  annotate("text", x=3, y=0.2, label="z = 2.13", size =8)
```

## Finding z from a given area

```{r echo=FALSE}
qplot(x1,y1,normDF, geom="line") + 
  geom_ribbon(data=subset(normDF, x1 >= -0.94), aes(ymax=y1), ymin=0, fill="blue",
              colour=NA, alpha=0.5) +
  geom_segment(aes(x=-0.94, y=0, xend=-0.94, yend=dnorm(-0.94))) +
  xlab("z") + ylab("") + ggtitle("Standard Normal Distribution") +
  annotate("text", x=0, y=0.1, label="0.8264", size=8)
```

## Finding z from a given area

```{r echo=FALSE}
qplot(x1,y1,normDF, geom="line") + 
  geom_ribbon(data=subset(normDF, x1 >= -0.94), aes(ymax=y1), ymin=0, fill="blue",
              colour=NA, alpha=0.5) +
  geom_segment(aes(x=-0.94, y=0, xend=-0.94, yend=dnorm(-0.94))) +
  xlab("z") + ylab("") + ggtitle("Standard Normal Distribution") +
  annotate("text", x=0, y=0.1, label="0.8264", size=8) +
  annotate("text", x=-2.7, y=0.25, label="1 - 0.8264 = -0.1736\nz = -0.94", size=7)
```

## Finding z from a given area

```{r echo=FALSE}
qplot(x1,y1,normDF, geom="line") + 
  geom_ribbon(data=subset(normDF, x1 >= 0.73), aes(ymax=y1), ymin=0, fill="blue",
              colour=NA, alpha=0.5) +
  geom_segment(aes(x=0.73, y=0, xend=0.73, yend=dnorm(0.73))) +
  xlab("z") + ylab("") + ggtitle("Standard Normal Distribution") +
  annotate("text", x=1.4, y=0.05, label="0.2327", size=6)
```

## Finding z from a given area

```{r echo=FALSE}
qplot(x1,y1,normDF, geom="line") + 
  geom_ribbon(data=subset(normDF, x1 >= 0.73), aes(ymax=y1), ymin=0, fill="blue",
              colour=NA, alpha=0.5) +
  geom_segment(aes(x=0.73, y=0, xend=0.73, yend=dnorm(0.73))) +
  xlab("z") + ylab("") + ggtitle("Standard Normal Distribution") +
  annotate("text", x=1.4, y=0.05, label="0.2327", size=6) +
  annotate("text", x=2.7, y=0.25, label="1 - 0.2327 = 0.7673\nz = 0.73", size=7)
```

## Finding z from a given area

```{r echo=FALSE}
qplot(x1,y1,normDF, geom="line") + 
  geom_ribbon(data=subset(normDF, x1 <= -1.28), aes(ymax=y1), ymin=0, fill="blue",
              colour=NA, alpha=0.5) +
  geom_segment(aes(x=-1.28, y=0, xend=-1.28, yend=dnorm(-1.28))) +
  xlab("z") + ylab("") + ggtitle("Standard Normal Distribution") +
  annotate("text", x=-1.7, y=0.04, label="0.1", size=7)
```

## Finding z from a given area

```{r echo=FALSE}
qplot(x1,y1,normDF, geom="line") + 
  geom_ribbon(data=subset(normDF, x1 <= -1.28), aes(ymax=y1), ymin=0, fill="blue",
              colour=NA, alpha=0.5) +
  geom_segment(aes(x=-1.28, y=0, xend=-1.28, yend=dnorm(-1.28))) +
  xlab("z") + ylab("") + ggtitle("Standard Normal Distribution") +
  annotate("text", x=-1.7, y=0.04, label="0.1", size=7) +
  annotate("text", x=-2.7, y=0.25, label="z = -1.28", size=8)
```

## Finding z from a given area

```{r echo=FALSE}
qplot(x1,y1,normDF, geom="line") + 
  geom_ribbon(data=subset(normDF, x1 >= 0.67), aes(ymax=y1), ymin=0, fill="blue",
              colour=NA, alpha=0.5) +
  geom_segment(aes(x=0.67, y=0, xend=0.67, yend=dnorm(0.67))) +
  xlab("z") + ylab("") + ggtitle("Standard Normal Distribution") +
  annotate("text", x=1.5, y=0.04, label="0.25", size=7)
```

## Finding z from a given area

```{r echo=FALSE}
qplot(x1,y1,normDF, geom="line") + 
  geom_ribbon(data=subset(normDF, x1 >= 0.67), aes(ymax=y1), ymin=0, fill="blue",
              colour=NA, alpha=0.5) +
  geom_segment(aes(x=0.67, y=0, xend=0.67, yend=dnorm(0.67))) +
  xlab("z") + ylab("") + ggtitle("Standard Normal Distribution") +
  annotate("text", x=1.5, y=0.04, label="0.25", size=7) +
  annotate("text", x=2.7, y=0.25, label="1 - 0.25 = 0.75\nz = 0.67", size=8)
```

## Standardizing

- Changes any normal model to a standard normal model
$$z = \frac{y - \mu}{\sigma}$$
- $z =$ # of standard deviations away from mean $\mu$
    - Negative $z\Rightarrow$ number is below mean
    - Positive $z\Rightarrow$ number is above mean

## Example

- The height of men follows  a normal distribution with mean 70 inches and standard deviation 3 inches
    - Standardize $y=68$ $$z = \frac{68-70}{3} = -0.67$$
    - Standardize $y=74$ $$z = \frac{74-70}{3} = 1.33$$
    
## Working with the normal model

- Finding areas

1. Standardize $y$ to get $z$
2. Find percentile for $z$ from Z-table
3. Use percentile to obtain needed area

## Working with the normal model

- Finding normal values

1. Determine percentile from given information in problem
2. Find $z$ value for percentile from Z-table
3. Calculate normal value $y$ using equation $$y = \mu + z\sigma$$

## Example 1A

- Height of men $\sim N(70,3)$
- What proportion of men are shorter than 66 inches?

$$z = \frac{66 - 70}{3} = -1.33$$

## Example 1A

```{r echo=FALSE}
male_height + 
  geom_ribbon(data=subset(mHeights, x <= 66), aes(ymax=y), ymin=0, fill="blue",
              colour=NA, alpha=0.5) +
  geom_segment(aes(x=66, y=0, xend=66, yend=dnorm(66, 70, 3))) +
  annotate("text", x=63, y=0.05, label="P(z<-1.33)\n=0.0918", size=7)
```

## Example 1B

- Height of men $\sim N(70,3)$
- What proportion of men are taller than 74 inches?
$$z = \frac{74 - 70}{3} = 1.33$$

## Example 1B

```{r echo=FALSE}
male_height + 
  geom_ribbon(data=subset(mHeights, x >= 74), aes(ymax=y), ymin=0, fill="blue",
              colour=NA, alpha=0.5) +
  geom_segment(aes(x=74, y=0, xend=74, yend=dnorm(74, 70, 3))) +
  annotate("text", x=77, y=0.08, label="P(z>1.33)\n=P(z<-1.33)\n=0.0918", size=8)
```

## Example 1C

- Height of men $\sim N(70,3)$
- What proportion of men are between 68 and 71 inches tall?
$$z_{1} = \frac{68 - 70}{3} = -0.67$$
$$z_{2} = \frac{71 - 70}{3} = 0.33$$

## Example 1C

```{r echo=FALSE}
male_height + 
  geom_ribbon(data=subset(mHeights, x >= 68 & x <= 71), aes(ymax=y), ymin=0, fill="blue",
              colour=NA, alpha=0.5) +
  geom_segment(aes(x=68, y=0, xend=68, yend=dnorm(68, 70, 3))) +
  geom_segment(aes(x=71, y=0, xend=71, yend=dnorm(71, 70, 3))) +
  annotate("text", x=63, y=0.08, label="P(z < -0.67)\n=0.2514", size=8) +
  annotate("text", x=63, y=0.05, label="P(z < 0.33)\n=0.6293", size=8) +
  annotate("text", x=77, y=0.08, label="0.6293 - 0.2514\n=0.3779", size=8)
```

## Example 1D

- The height of men is known to follow a normal model with mean 70 and standard deviation 3
- 20% of men are taller than what height?
    - Look up 0.8 in the table and work backwards to find $z$
    - $z = 0.84$
    - $y = 70 + 0.84(3) = 72.52$
    
## Example 1E

- The height of men is known to follow a normal model with mean 70 and standard deviation 3
- 65% of men are shorter than what height?
    - Look up 0.65 in the table and work backwards to find $z$
    - $z = 0.39$
    - $y = 70 + 0.39(3) = 71.17$
    
## Unusual?

- The height of men is known to follow a normal model with mean 0 and standard deviation 3
- Would it be unusual to find a man that is more than 80 inches tall? Why or why not?
    - His z-score is $3.33$. He is more than 3 standard deviations above the mean. So yes, that is unusual
    - The probability of being 80" tall or more is 0.0004
- Would it be unusual to find a man that is less than 59 inches tall? Why or why not?
    - His z-score is -$3.67$. He is more than 3 standard deviations below the mean. So yes, that is unusual
    - The probability of being less than or equal to 59" tall is 0.0001
    
## Unusual?

- If you have an observation that is more than 3 standard deviations from the mean (z-score more than 3 or less than -3), it is considered to be an unusual occurence

